Type Refinements for the Working Class

### 2015-12-29 paper.tex
remove aborted sentence
Update notation package

\newcommand\EvalN[5]{\IMode{#1}\parallel\IMode{#4}\Downarrow^{\OMode{#3}}_{\IMode{#2}}\OMode{#5}}
\newcommand\Eval[4]{\EvalN{#1}{#2}{}{#3}{#4}}
\newcommand\Exprs{\mathbf{E}}
\newcommand\Values{\mathbf{V}}
\newcommand\BTms{\mathbf{B}}
\newcommand\Naturals{\mathbb{N}}
\newcommand\NaturalsPsh{\mathcal{N}}

@@

For a lazy computation language
$\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define
sort-indexed (resp.\ valence-indexed) copresheaves on $\SCtx$ of closed values
(resp.\ bound terms) as follows:
\begin{align*}
  \ADefine{\Values_\tau(\Upsilon)}{
    \MkSet{
      M\equiv\App{\vartheta}{\vec{E}}
      \mid M\in\Exprs_\tau(\Upsilon)\land \exists a. \CanOperators_a(\Upsilon)\ni\vartheta
    }
  }\\
  \ADefine{
    \BTms_{v}(\Upsilon)
  }{
    \MkSet{E \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\cdot}{E}{v}}
  }
\end{align*}

@@

Next, we can always extend $R_\tau$ to a new
family of valenced-indexed relations
$\Of{R_v}{\Hom{\SemU}{\mathbb{1}}{\Pow{\BTms_v^2}}}$ for
$\Match{v}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$, defined pointwise as
follows:
\[
  \infer{
    \PBinRel{R_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}}{\Upsilon}{
      \MkBTm{\vec{u}}{\vec{x}}{M}
    }{
      \MkBTm{\vec{v}}{\vec{y}}{N}
    }
  }{
    \begin{array}{l}
      \forall\IMode{\vec{w}}\mathrel{\#}\IMode{\Dom{\Upsilon}\cup\vec{u}\cup\vec{v}}.\quad
      \forall\IMode{\vec{z}}\mathrel{\#}\IMode{\vec{x}\cup\vec{y}}.\\
      \HPBinRel{R_\tau}{\Upsilon,\vec{w}:\vec{\sigma}}{\vec{z}:\vec{\tau}}{
        \Subst{\vec{z}}{\vec{x}}{
          \Rename{(\vec{u}\mapsto\vec{w})}{M}
        }
      }{
        \Subst{\vec{z}}{\vec{y}}{
          \Rename{(\vec{v}\mapsto\vec{w})}{M}
        }
      }
    \end{array}
  }
\]

@@

  An operator $\IsOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}}$ is
  \emph{extensional} when, for any
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}%
  , for all $\Member{\vec{E},\vec{F}}{\BTms^{[\vec{v}]}(\Upsilon')}$ and
  $\Member{M}{\Exprs_\tau(\Upsilon')}$ and $\IMode{k}\geq\IMode{0}$, if we can conclude
  $\PBinRel{\preccurlyeq^\star_\tau}{\Upsilon'}{M}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$
  from the following conditions:
  %
  \begin{enumerate}
    \item
      We have $\EvalN{\Upsilon'}{\tau}{k}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{M}$.

@@

\begin{proof}
  Fix
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}%
  and $\Member{\vec{E},\vec{F}}{\BTms^{[\vec{v}]}(\Upsilon')}$ and
  $\Member{M}{\Exprs_\tau(\Upsilon')}$ and $\IMode{k}\geq\IMode{0}$.
  %
  By hypothesis (1) from Definition~\ref{def:extensionality} above, we have
  $\EvalN{\Upsilon'}{\tau}{k}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{M}$;
  because $M$ is a closed term, by the idempotency of evaluation we have
  $\Match{M}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}$.

@@

We will demonstrate how to define certain refinements, and prove that they are
types. The refinements below will differ from the ones defined in Section~

>>>>

\newcommand\EvalN[5]{\IMode{#1}\parallel\IMode{#4}\Downarrow^{\OMode{#3}}_{\IMode{#2}}\OMode{#5}}
\newcommand\Eval[4]{\EvalN{#1}{#2}{}{#3}{#4}}
\newcommand\Exprs{\mathbf{E}}
\newcommand\Values{\mathbf{V}}
\newcommand\BExprs{\mathbf{B}}
\newcommand\Naturals{\mathbb{N}}
\newcommand\NaturalsPsh{\mathcal{N}}

@@

For a lazy computation language
$\Match{L}{\Tuple{\Sorts,\Operators,\CanOperators}}$, let us define
sort-indexed (resp.\ valence-indexed) copresheaves on $\SCtx$ of closed values
(resp.\ bound terms) as follows:
\begin{align*}
  \ADefine{\Values_\tau(\Upsilon)}{
    \MkSet{
      M\equiv\App{\vartheta}{\vec{E}}
      \mid M\in\Exprs_\tau(\Upsilon)\land \exists a. \CanOperators_a(\Upsilon)\ni\vartheta
    }
  }\\
  \ADefine{
    \BExprs_{v}(\Upsilon)
  }{
    \MkSet{E \mid \IsAbtUnmoded{\cdot}{\Upsilon}{\cdot}{E}{v}}
  }
\end{align*}

@@

Next, we can always extend $R_\tau$ to a new
family of valenced-indexed relations
$\Of{R_v}{\Hom{\SemU}{\mathbb{1}}{\Pow{\BExprs_v^2}}}$ for
$\Match{v}{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}$, defined pointwise as
follows:
\[
  \infer{
    \PBinRel{R_{\MkValence{\vec{\sigma}}{\vec{\tau}}{\tau}}}{\Upsilon}{
      \MkAbs{\vec{u}}{\vec{x}}{M}
    }{
      \MkAbs{\vec{v}}{\vec{y}}{N}
    }
  }{
    \begin{array}{l}
      \forall\IMode{\vec{w}}\mathrel{\#}\IMode{\Dom{\Upsilon}\cup\vec{u}\cup\vec{v}}.\quad
      \forall\IMode{\vec{z}}\mathrel{\#}\IMode{\vec{x}\cup\vec{y}}.\\
      \HPBinRel{R_\tau}{\Upsilon,\vec{w}:\vec{\sigma}}{\vec{z}:\vec{\tau}}{
        \Subst{\vec{z}}{\vec{x}}{
          \Rename{(\vec{u}\mapsto\vec{w})}{M}
        }
      }{
        \Subst{\vec{z}}{\vec{y}}{
          \Rename{(\vec{v}\mapsto\vec{w})}{M}
        }
      }
    \end{array}
  }
\]

@@

  An operator $\IsOperator{\Upsilon}{\vartheta}{\MkArity{\vec{v}}{\tau}}$ is
  \emph{extensional} when, for any
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}%
  , for all $\Member{\vec{E},\vec{F}}{\BExprs^{[\vec{v}]}(\Upsilon')}$ and
  $\Member{M}{\Exprs_\tau(\Upsilon')}$ and $\IMode{k}\geq\IMode{0}$, if we can conclude
  $\PBinRel{\preccurlyeq^\star_\tau}{\Upsilon'}{M}{\App{\Rename{\rho}{\vartheta}}{\vec{F}}}$
  from the following conditions:
  %
  \begin{enumerate}
    \item
      We have $\EvalN{\Upsilon'}{\tau}{k}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{M}$.

@@

\begin{proof}
  Fix
  \begin{tikzcd}[cramped]
    \IMode{\Upsilon} \arrow[r,hook,"\IMode{\rho}"] &\IMode{\Upsilon'}
  \end{tikzcd}%
  and $\Member{\vec{E},\vec{F}}{\BExprs^{[\vec{v}]}(\Upsilon')}$ and
  $\Member{M}{\Exprs_\tau(\Upsilon')}$ and $\IMode{k}\geq\IMode{0}$.
  %
  By hypothesis (1) from Definition~\ref{def:extensionality} above, we have
  $\EvalN{\Upsilon'}{\tau}{k}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}{M}$;
  because $M$ is a closed term, by the idempotency of evaluation we have
  $\Match{M}{\App{\Rename{\rho}{\vartheta}}{\vec{E}}}$.

@@

We will demonstrate how to define certain refinements, and prove that they are
types.
